---
title: "h4"
author: "Yating Tian"
date: "2/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. For each of parts (a) through (d), indicate whether we would generally
expect the performance of a flexible statistical learning method to be
better or worse than an inflexible method. Justify your answer.
(a) The sample size n is extremely large, and the number of predictors p is small.
    
    **The flexible statistical learning method is better because it will fit the data closer and better when the large sample size.**
    
(b) The number of predictors p is extremely large, and the number of observations n is small.

    ** A flexible statistical learning method is worth because it would overfit when there is the small n**
    
(c) The relationship between the predictors and response is highly non-linear.

    **the flexible statistical learning method is better because it will fit the data closer than inflexible method.

(d) The variance of the error terms, i.e. σ2 = Var(), is extremely
high.

    ** A flexible statistical learning method is worth because it would fit the noise too, which brings more varience to the prediction. **
 
2. Explain whether each scenario is a classification or regression problem, and indicate whether we are most interested in inference or prediction. Finally, provide n and p.

(a) We collect a set of data on the top 500 firms in the US. For each firm we record profit, number of employees, industry and the CEO salary. We are interested in understanding which factors affect CEO salary.

    **regression; inference; n=500;p= 3, (profit, number of employees, industry)**
    
(b) We are considering launching a new product and wish to know whether it will be a success or a failure. We collect data on 20 similar products that were previously launched. For each product we have recorded whether it was a success or failure, price charged for the product, marketing budget, competition price, and ten other variables.

    **classification;prediction; n=20; p=13,(price charged for the product, marketing budget, competition price, and ten other variables)**
    
(c) We are interest in predicting the % change in the USD/Euro exchange rate in relation to the weekly changes in the world stock markets. Hence we collect weekly data for all of 2012. For each week we record the % change in the USD/Euro, the %
change in the US market, the % change in the British market, and the % change in the German market.
    
    **regression;prediction;n= 52;p= 3,(the % change in the US market, the % change in the British market, and the % change in the German market.) **



This exercise involves the Boston housing data set.

```{r}

#(a) To begin, load in the Boston data set. The Boston data set is part of the MASS library in R. Now the data set is contained in the object Boston. How many rows are in this data set? How many columns? What do the rows and columns represent?
library(tidyverse)
library(MASS)
library(reshape2)
view(Boston)
nrow(Boston) #This is the number of obervation
ncol(Boston) #This is the number of measured varables of each overvation

#(b) Make some pairwise scatterplots of the predictors (columns) in this data set. Describe your findings.

plot(Boston$indus, Boston$tax)   #By observing graph, the relationship proportion of non-retail business acres per town and tax has weak positive correlation.
plot(Boston$nox, Boston$tax) #the ralationship between nitrogen oxides concentration and tax is that for more than 0.65 parts per 10 million, more likely to have higher full-value property-tax rate per \$10,000. 
plot(Boston$age, Boston$tax) #people with higher age(after 40) more likely to have highter full-value property-tax rate per \$10,000.  


#(c) Are any of the predictors associated with per capita crime rate? If so, explain the relationship.
hist(Boston$crim, breaks = 100) #most of town doesn't have crime or happen 0-1 per capita crime rate. Over 90 % have crime less than 20. 

bos <- melt(Boston, id="crim")
ggplot(bos, aes(x=value, y=crim)) +
  facet_wrap(~variable, scales="free") + 
  geom_point()
#there is no strong liner relationship beween each predictors and crime rate. age/dis/balck/medv might have some non-liner relationship.

#(d) Do any of the suburbs of Boston appear to have particularly high crime rates? Tax rates? Pupil-teacher ratios? Comment on the range of each predictor.
max(Boston$crim)
Boston[Boston$crim == max(Boston$crim),]
range(Boston$crim) # 0.00632 88.97620
Boston[Boston$tax == max(Boston$tax),]
range(Boston$tax) #187 711
max(Boston$ptratio)
Boston[Boston$ptratio == max(Boston$ptratio),]
range(Boston$ptratio)  #12.6 22.0

boxplot(Boston$crim)

#There is many outliers has particularly high crime rate. 

boxplot(Boston$tax)

#There is no outliers has particularly high tax rate. 

boxplot(Boston$ptratio) 

# There is no outliers has particularly high Pupil-teacher ratios. There is only couple outliers that has particularly low Pupil-teacher ratios

#(e) How many of the suburbs in this data set bound the Charles river?
nrow(Boston[Boston$chas == 1, ])

#(f) What is the median pupil-teacher ratio among the towns in this data set?
median(Boston$ptratio)

#(g) Which suburb of Boston has lowest median value of owneroccupied homes? What are the values of the other predictors for that suburb, and how do those values compare to the overall ranges for those predictors? Comment on your findings.

Boston[min(Boston$medv),]

range(Boston$crim)
Boston[min(Boston$medv),1] 
#the suburb have really low crime rate compare to the range of crime rate that 0.00632 88.97620. 

range(Boston$zn)
Boston[min(Boston$medv),2] 
#the suburb have really low proportion of residential land zoned for lots over 25,000 sq.ft.,which is 0 compare to the over all range 0 to 100

range(Boston$indus)
Boston[min(Boston$medv),3] 
#the suburb have reletively low proportion of non-retail business acres per town 2.18 compare to the over all range 0.46 to 27.74

range(Boston$chas)
Boston[min(Boston$medv),4] 
# Charles River dummy variable (= 1 if tract bounds river; 0 otherwise). so this suburb has otherwise river.

range(Boston$nox)
Boston[min(Boston$medv),5] 
#the suburb have reletively low nitrogen oxides concentration (parts per 10 million). which is 0.458 compare to the over all range 0.385 to 0.871

range(Boston$rm)
Boston[min(Boston$medv),6] 
#the suburb have reletively high average number of rooms per dwelling. which is 7.147 compare to the over all range 3.561 to 8.780

range(Boston$age)
Boston[min(Boston$medv),7] 
# age is about the middle comapre to overall range 

range(Boston$dis)
Boston[min(Boston$medv),8] 
#the suburb's wighted mean of distances to five Boston employment centres is at middle, which is 6.0622 compare to the over all range 1.1296 12.1265

range(Boston$rad)
Boston[min(Boston$medv),9] 
#the suburb have reletively low index of accessibility to radial highways, which is 3 compare to the over all range 1 to 24. 

range(Boston$tax)
Boston[min(Boston$medv),10] 
# this suburb's full-value property-tax rate per \$10,000 was relitively low.

range(Boston$ptratio)
Boston[min(Boston$medv),11] 
#the suburb have reletively middle pupil-teacher ratio by town, which is 18.7 compare to the over all range 12.6 to 22.0

range(Boston$black)
Boston[min(Boston$medv),12] 
#the suburb have the highest 1000(Bk - 0.63)^2 where Bk is the proportion of blacks by town.1000(Bk - 0.63)^2 where Bk is the proportion of blacks by town. which is 396.9 compare to the over all range 0.32 to 396.90

range(Boston$lstat)
Boston[min(Boston$medv),13]
# 5.33 ower status of the population (percent) compare to overall range 1.73 37.97, which is at the lower level. 

 

#(h) In this data set, how many of the suburbs average more than seven rooms per dwelling? More than eight rooms per dwelling? Comment on the suburbs that average more than eight rooms per dwelling.

nrow(Boston[Boston$rm > 7, ])
nrow(Boston[Boston$rm > 8, ]) 

# there was 13 suburbs have more than eight rooms per dwelling. 
```

