---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE)
library(tidyverse)
library(lubridate)
```

```{r viridis-default, include=FALSE}
## reset color defaults
## Source https://data-se.netlify.com/2018/12/12/changing-the-default-color-scheme-in-ggplot2/
library(viridis)
library(scales)

#### continuous variables color and fill
options(ggplot2.continuous.colour = "viridis")
options(ggplot2.continuous.fill = "viridis")

#### use viridis for discrete scales
scale_colour_discrete <- scale_colour_viridis_d
scale_fill_discrete <- scale_colour_viridis_d

## reset default theme
##theme_set(theme_minimal())
```

## Assignment 5

### Yating Tian
### netid: ytian83

#### Due Thursday, October 10

The purpose of this assignment is to give you practice using tidyr commands and to review dplyr, ggplot2, and basic exploratory data analysis skills.

Turn in an HTML file and this R Markdown file after you have edited it.

### Data

The data set is in the file `ecology.csv`.
It contains records about animals that were caught in traps
during a many-year ecological experiment of a desert ecosystem near Portal, Arizona.

```{r read-data, include=FALSE}
eco = read_csv("ecology.csv")
```

### Problems

### 1

> The primary scientific objectives related to rodents. However, the traps also caught other species.
Filter the data set to eliminate all observations where the `taxa` variable is not Rodent. How many cases remain?

```{r}
eco=eco %>%
  filter(taxa == "Rodent")

#34786 obs to 34247 obs

```

### 2

> The data set has a variable `species_id` which is a two-letter code for each species. It also has columns named `genus` and `species`. The conventional name of a species is to state both the genus and species separated by a space. Add to the data set a column named `species_name` that contains the `genus` and `species` values separated by a space.
Then summarize the data to count how many rodent observations there are by species name, and arrange from most to least.
How many distinct species are there?
Which appears most often and how many of these individuals are in the data set?

```{r}
eco=eco %>%
  unite("species_name","genus","species",sep = " ")
  
eco %>% 
  group_by(species_name) %>%
  summarise(n=n()) %>%
  arrange(desc(n))

##29 distinct species
##Dipodomys merriami appears most
```

### 3

> Data was collected over many years and in many plots. Make a table with one row for each plot and one column for each year with values equal to the count of the number of rodent observations collected in each plot each year with zeros where appropriate. (Use `select()`, followed by `group_by()` and `summarize()` or `count()` followed by `spread()`.)

```{r}
eco %>%
  select(year,plot_id)%>%
  group_by(plot_id,year)%>%
  summarise(n=n())%>%
  spread(key=year,value=n)

```

### 4

> What is the mean weight of each rodent species and plot? Drop any cases with missing weight information. Display the information in a table with a row for each species and a column for each plot. Combinations without data should have the missing data symbol in the appropriate cell in this table.
What are the dimensions of this table?
Which species/plot has the highest average weight? The lowest?

```{r}
eco %>%
  select(species_name,plot_id,weight)%>%
  filter(is.na(weight)==FALSE) %>%
  group_by(species_name,plot_id)%>%
  summarise(mean_weight=mean(weight))%>%
  spread(key=plot_id, value = mean_weight) 

#this table is a two demention table, which contains both species name and plot id 
  
eco %>%
  select(species_name, weight, plot_id) %>%
  group_by(species_name,plot_id) %>%
  summarise(mean=mean(weight,na.rm = TRUE))%>%
  top_n(n=1, wt=mean)%>%
  arrange(desc(mean))

#Sigmodon fulviventer	in 23	has a highest weight which is 199.000000	

eco %>%
  select(species_name, weight, plot_id) %>%
  group_by(species_name,plot_id) %>%
  summarise(mean=mean(weight,na.rm = TRUE))%>%
  top_n(n=-1, wt=mean)%>%
  arrange((mean))

#Baiomys taylori in plot 2	has the lowest weight, which is 6.000000	

```

### 5

> Use the wasp data from discussion. Correct the error. Then use `gather()` to rewrite the data so that there is a single column of count and a separate column for the species of wasp that was counted.
Then summarize this long data to find the total number of wasps collected in traps by species.
Arrange these counts in a table from highest to lowest counts.

```{r}
library(readxl)
wasps = read_excel("wasps-data.xlsx") %>%
  rename(date = `Date Collected`,
         block = Block,
         position = Number,
         lure = Lure,
         maculifrons = `V. Maculifrons`,
         flavopilosa = `V. Flavopilosa`,
         germanica = `V. Germanica`,
         vidua = `V. Vidua`) %>%
  mutate(date = as_date(date)) %>%
  mutate(week = as.integer(as.factor(date))) %>%
  mutate(two_mb = as.integer(str_detect(lure,"2MB"))) %>%
  mutate(aa = as.integer(str_detect(lure,"AA"))) %>%
  mutate(ib = as.integer(str_detect(lure,"IB"))) %>%
  select(date,week,everything())

wasps$position[160]=2
wasps$block[160]= "D"
```

```{r}
wasps1=wasps %>%
  gather(key="species", value = "count", maculifrons:vidua)

wasps1 %>%  
  group_by(species)%>%
  summarise(total=sum(count))%>%
  arrange(desc(total))

```

### 6

> Make a plot that shows the number of wasps collected by species in a box and whisker plot with jittered points overlayed. Use faceting so that each week is in a different row.

```{r}
wasps1%>%
  ggplot(aes(x=species,y=count,color=species))+
  geom_boxplot()+
  geom_point(position=position_jitter(height = 0,width=0.1))+
  facet_wrap(.~week)
```

